<!-- R Commander Markdown Template -->

Homework #3
=======================

### Taylor Thul

### `r as.character(Sys.Date())`

```{r echo=FALSE}
# include this code chunk as-is to set options
knitr::opts_chunk$set(comment=NA, prompt=TRUE, out.width=750, fig.height=8, fig.width=8)
library(Rcmdr)
library(car)
library(RcmdrMisc)
```


```{r echo=FALSE}
# include this code chunk as-is to enable 3D graphs
library(rgl)
knitr::knit_hooks$set(webgl = hook_webgl)
```

Get the dataset into R.

```{r}
names(White_wines) <- make.names(names(White_wines))
```

```{r}
library(abind, pos=18)
```


```{r}
library(e1071, pos=19)
```

Run some summary statistics: 

```{r}
numSummary(White_wines[,c("alcohol", "chlorides", "citric.acid", "density", 
  "fixed.acidity", "free.sulfur.dioxide", "pH", "quality", "residual.sugar", 
  "sulphates", "total.sulfur.dioxide", "volatile.acidity")], 
  statistics=c("mean", "sd", "IQR", "quantiles"), quantiles=c(0,.25,.5,.75,1))
```
Make sure no data is missing:

```{r}
sapply(White_wines, function(x)(sum(is.na(x)))) # NA counts
```
No missing data! Hooray!
Look at the distribution of the variables:

```{r}
#appears normal
with(White_wines, Hist(quality, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```

Scatterplot matric of alcohol, chlorides, citric acid and quality
```{r}
scatterplotMatrix(~alcohol+chlorides+citric.acid+quality, reg.line=FALSE, 
  smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, levels=c(.5, .9), 
  id.n=0, diagonal = 'density', data=White_wines)
```

Scatterplot of density, fixed acidity, free sulfur dioxide and pH

```{r}
scatterplotMatrix(~density+fixed.acidity+free.sulfur.dioxide+pH, 
  reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```

Scatterplot of residual sugar, sulphates, total sulfur dioxide and volatile acidity.  Residual sugar may need transformed. 

```{r}
scatterplotMatrix(~residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   reg.line=FALSE, smooth=FALSE, spread=FALSE, span=0.5, ellipse=FALSE, 
  levels=c(.5, .9), id.n=0, diagonal = 'density', data=White_wines)
```
look at residual sugar on it's own

```{r}
with(White_wines, Hist(residual.sugar, scale="frequency", breaks="Sturges", 
  col="darkgray"))
```
add new variable of log of sugar residuals
```{r}
White_wines$residual.sugar2 <- with(White_wines, log2(residual.sugar))
```
it doesn't really look more normal 

```{r}
with(White_wines, Hist(residual.sugar2, scale="frequency", breaks="Sturges",
   col="darkgray"))
```

```{r}
library(nortest, pos=20)
```

```{r}
with(White_wines, shapiro.test(residual.sugar))
```
the Shapiro-Wilk is < 0.05 so we accept the hypothesis the data is not normally distibuted.
so try the Shapiro- Wilk on the transformed data 

```{r}
with(White_wines, shapiro.test(residual.sugar2))
```
this did not help at all... still not normal.
moving on to the regression model


....but I don't know what to do about that so moving on the the regression
```{r}
# divide the dataset into a training and a testing set based on a random uniform number on fixed seed, which in this case we are using the date
# this step is also creating a new variable and adding it to the data set which is a distribution of random numbers from 0 to 1 

set.seed(20170214)
White_wines$group <- runif(length(White_wines$quality), min = 0, max = 1)
summary(White_wines)


#what random forests do is this process over and over again and makes the aggregate which might be called bootstrapping?

White_wines.train <- subset(White_wines, group <= 0.90)
White_wines.test <- subset(White_wines, group > 0.90)

#see if it worked
summary(White_wines.train)

#I think it worked! 
```
```

moving on to do a regression:


```{r}
RegModel.1 <- 
  lm(quality~alcohol+chlorides+citric.acid+density+fixed.acidity+free.sulfur.dioxide+pH+residual.sugar+sulphates+total.sulfur.dioxide+volatile.acidity,
   data=White_wines.train)
summary(RegModel.1)
```
So it looks like alcohol, chlorides, density, fixed acidity, pH, residual. sugar, sulphates, and volatile acidity are all singificanly assoicated with predicting quality.  
Run the model again taking out the non-signficant variables:

```{r}
RegModel.2 <- 
  lm(quality~alcohol+density+fixed.acidity+free.sulfur.dioxide+pH+residual.sugar+sulphates++volatile.acidity,
   data=White_wines.train)
summary(RegModel.2)
```
Check some plots to see how the model fits:
```{r}
layout(matrix(1:4,2,2))
plot(RegModel.2)
```


